\sysfull

### 2018-08-04 paper/asplos19/background.tex
WAR figure updated

A power failure clears volatile state (e.g., registers and SRAM) while
non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
to a prior point in the execution: by default, to the beginning of {\tt
main()}. Early intermittent systems preserved progress by periodically
checkpointing volatile execution context to non-volatile
memory~\cite{mementos}, sometimes using hardware
support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. 
%
Checkpointing volatile state alone does not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Precisely, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
Figure~\ref{fig:war} illustrates how state can become
inconsistent in an intermittent execution using a simple example of an average
operation over an array of integers.
The variable \texttt{sum} introduces the WAR, being read and written sequentially
by the increment operation, thus being erroneously incremented when a power
failure occurs before updating the index \texttt{i}.

>>>>

A power failure clears volatile state (e.g., registers and SRAM) while
non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
to a prior point in the execution: by default, to the beginning of {\tt
main()}. Early intermittent systems preserved progress by periodically
checkpointing volatile execution context to non-volatile
memory~\cite{mementos}, sometimes using hardware
support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. 
%
Checkpointing volatile state alone does not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Precisely, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
Figure~\ref{fig:war} illustrates how state can become
inconsistent in an intermittent execution using a simple example of an average
operation over an array of integers.
The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially
by the increment operation.
If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).
